---
title: "HICP Forecasting for Norway (2010-2021)"
output: html_notebook
author : "Renuka Nair"
---

Loading the data for HICP of basket goods (sugar, honey, jam and other confectionary)in Norway 2010-2021

```{r}
library(fpp)
library(fpp2)
library(TTR)

HICP_Norway <- read.csv("C:/Users/Renuka/Downloads/HICP_Norway.csv")
```

```{r}
class(HICP_Norway)
```

Converting the dataset into a time series for forecasting 

```{r}
hicp_ts <- ts (HICP_Norway$HICP, frequency=12, start=c(2010,1))
hicp_ts
plot (hicp_ts, main="HICP for Basket Goods (Sugar,Honey,Jam,Confectionery) in Norway(2010-2021)")

```
Summary of the time series data :

```{r}
summary(hicp_ts)
```

## Forecasting

### Mean Forecast
We start with the simplest of forecast first which is mean forecast that simply takes the average of the observations and forecasts the value 

```{r}
mean_forecast <- meanf(hicp_ts,5)
plot (mean_forecast)
```

### Naive Forecast
Naive Forecasting principle considers the latest data points to be most relevant 

```{r}
naive_forecast <- naive(hicp_ts,5)
plot(naive_forecast)
lines(naive_forecast$mean,col="yellow")
```
### Random Walk Forecast
Similar to earlier forecasts, this is a low cost option 
```{r}
rwf_forecast <- rwf(hicp_ts,5)
plot(rwf_forecast)
```

### Seasonal Naive Forecast
Seasonal Naive takes the effects of seasonality into account while forecasting

```{r}
snaive_forecast <- snaive(hicp_ts,5)
plot(snaive_forecast)
lines(snaive_forecast$mean,col="yellow")
lines(snaive_forecast$fitted,col="purple")
```

We can see that seasonal naive is not that great at forecasting as the error in forecasting seems quite huge at places.

### Exponential Smoothing
Exponential smoothing tells us whether the 
```{r}
ets_fit <- ets(hicp_ts)
summary(ets_fit)
plot(ets_fit)

```

From the summary we can see that the data shows Multiplicative errors, No trend and Additive seasonality.

Mean Squared Error for ets :

```{r}
ets_fit$mse
```

Forecasting can be done by :
```{r}
forecast_ets <- forecast.ets(ets_fit, h=5)
forecast_ets
plot(forecast_ets)
```

